---
title: "CentralDogmaBio: DNA Sequence Analysis Package Vignette"
output: rmarkdown::pdf_document
vignette: >
  %\VignetteIndexEntry{CentralDogmaBio}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Link to Git: <https://github.com/rforbiodatascience24/group_19_package>

The package CentralDogmaBio provides a set of tools to make work with DNA sequences easier. It allows users to create random DNA sequences, convert them into RNA sequences, extract codons, translate the codons into amino acids, and visualize amino acid counts in any given sequence.

```{r}
options(bitmapType = "cairo")
```

```{r}
library(CentralDogmaBio)
```

## Functions 

1.  create_DNA_string

Description: Generates a DNA string of random sequence with the input length.

2.  DNA_to_RNA

Description: Converts a DNA sequence to an RNA sequence.

3.  extract_codons

Description: Converts a nucleotide sequence into codons.

4.  translate_codons

Description: Translates codons into amino acids

5.  extract_char

Description: Analyzes the frequency of unique characters (amino acids) in a given sequence and creates a bar plot.

## Combining Functions

Here is a workflow that uses the functions from the package to go from DNA sequence to visualizing amino acid counts in the RNA translation:

```{r}
# Generate DNA sequence
dna_sequence <- create_DNA_string(12)

# Convert DNA to RNA
rna_sequence <- DNA_to_RNA(dna_sequence)

# Extract codons
codons <- extract_codons(rna_sequence)

# Translate codons to amino acids
amino_acid_sequence <- translate_codons(codons)
print(amino_acid_sequence)

# Visualize character counts in DNA sequence
extract_char(dna_sequence)
```

## Use and Future Additions

This package can be applied to:

-   Simulate DNA and RNA sequences for educational purposes

-   Perform basic analyses of sequences to understand base distributions

-   Translate RNA sequences to amino acid sequences (useful in basic bioinformatics)

Future versions could include:

-   More comprehensive error checking for invalid sequences

-   Add a function to generate the reverse complement of DNA (for DNA sequence analysis)

-   A function that can add mutations to the DNA you are working with to study mutations

-   A function that detects different mutation types
